---
title: "PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Steps Analysis

### Getting the code
```{r}
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip','data.zip')
unzip('data.zip')
activity_data <- read.csv('activity.csv')
head(activity_data)
```
Basic processing and summary stats
```{r}
activity_data$date <- as.Date(activity_data$date)
mean(is.na(activity_data$steps))
summary(activity_data)
```

### 1. What is mean total number of steps taken per day?

```{r}
steps_per_day_total <- with(activity_data, aggregate(steps, by=list(date), sum))

names(steps_per_day_total) <- c("date", 'steps')
head(steps_per_day_total)
plot(steps_per_day_total$steps ~ steps_per_day_total$date, main="Total Steps/Day", type="h", lwd=5, xlab="Date", ylab="Step Count")

```

### 2. What is the average daily activity pattern?

```{r}
stepsInterval<-aggregate(steps~interval,data=activity_data,mean,na.rm=TRUE)
plot(steps~interval,data=stepsInterval,type="l", main="Average Steps/Interval")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
stepsInterval[which.max(stepsInterval$steps), ]
```

### 3. Imputing missing values
```{r}
summary(steps_per_day_total)
NA2mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
imputed_data <- replace(activity_data, TRUE, lapply(activity_data, NA2mean))
head(imputed_data)
imputed_data_steps_per_day <- aggregate(imputed_data$steps, by=list(imputed_data$date), sum)
names(imputed_data_steps_per_day) <- c('date', 'steps')
head(imputed_data_steps_per_day)
plot(imputed_data_steps_per_day$steps ~ imputed_data_steps_per_day$date, main="(Imputed) Total Steps/Day", type="h", lwd=5, xlab="Date", ylab="Step Count")
```

Summary statistics of imputed data
```{r}
summary(imputed_data_steps_per_day)
```

### 4. Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#Use `%in%` and `weekdays` to create a logical vector
#convert to `factor` and specify the `levels/labels`
imputed_data$week_day <- factor((weekdays(imputed_data$date) %in% weekdays1), 
                   levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
aggregate_by_weekday <- aggregate(steps~interval+week_day, data=imputed_data, mean)
xyplot(steps ~ interval | week_day, data=aggregate_by_weekday, type="l")

```